---
title: "Osm Time Travel"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Osm Time Travel}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
This vignettes consists of 

* [1. Introduction](#ch1)
* [2. Attic Functions](#ch2)
* [3. Classes and methods](#ch3)
* [4. How to use](#ch4)
* [5. Future work](#ch5)

## 1. Introduction {ch1}
OSM Overpass Api supports query for attic data but this feature has not been implemented in R. This issue is discovered [here][1]. This package is made with intention of extending the [Osmdata][2] package with the attic functions. It has 2 classes to handle attic data `osmtt` and `osmttdiff`. Osmtt represents data from a previous timestamp and osmttdiff represents changes in data between the given time frame. 

## 2. Attic Functions {ch2}
[Attic settings][3] in overpass-api are *date,diff and adiff*. *date* returns results based on the OpenStreetMap database as of the date specified. *Diff* lets the database determine the difference of two queries at different points in time. *Adiff* is same like *diff*, but for all elements that aren't contained in the newer result, it is indicated what happened to them. In this package `add_date` and `add_adiff`, are implemented. *diff* is omitted since *adiff* gives more details than *diff*.

## 3. Classes and methods

### Class: 
### 1. osmtt

`osmtt` adds *datetime* parameters to `overpass_query` class which is then passed onto the actual query. The return from `overpass_query` and `osmtt` has the same structure so `osmtt` works similar to `overpass_query`. A **[bug][4]** in the `osmdata` package was found during this implementation. It has been fixed and pull request has been sent.

#### Methods 
* print: Print returns the overpass query string from the object, it can be passed to overpass-api from browser too.
* plot: Plots the points and lines from osmdata object
* osmtt_sf: queries overpass object and converts to osmdata object

### 2. osmttdiff
`osmttdiff` handles the diff results from adiff setting. It separates elements according to create, modify and delete actions. Also it separates the new and old ones.

#### Methods
* print: 
* osmttdiff_sf:

## 4. How to use
Attic data may not be supporeted by all overpass servers. First step is to find one which can handle the query.
```{r}
# new_url <- "http://overpass-api.de/api/interpreter"
# osmdata::set_overpass_url(new_url)
```
To get the instance of osm database at a previous version in time, Use `osmtt` object.
```{r}
# op <- osmtt("sudal nepal","2019-01-01T12:00:00Z")
```
Convert it to osmdata_sf object from `osmdata` package.
```{r}
# op_sf <- osmtt_sf(op)
```
Other formats can also be used by directly using the `osmdata` package methods
```{r}
# osmdata::osmdata_sp(op)
# osmdata::osmdata_xml(op)
```
(Note: Queries are commented because query quota was exceeded)

## 5. Issues, Challenges and Further Works 
### 5.1 Issues:
* When overpass api is busy, `osmttdiff` does not handle wait and retry.
* `osmtt` uses native call to C for parsing in `osmtt_sf`, but is not available in `osmttdiff`. Hence `osmttdiff` processing is slower.
* Writing `xml` output to file is supported but reading from `xml` has not been implemented

### 5.2 Challenges
* Bugs in the dependency gave unexpected results, which took much time to figure and solve.
* Running out of api query quota for overpass-api
* Some code were in cpp. Knowledge on this part was limited
* Documentation on adiff format is limited

### 5.3 Further Works
* Visualization of changes
* Integration of osmtt and osmttdiff for piping

  [1]: https://github.com/ropensci/osmdata/issues/179
  [2]: https://cran.r-project.org/web/packages/osmdata/index.html
  [3]: https://wiki.openstreetmap.org/wiki/Overpass_API/Overpass_QL#date
  [4]: https://github.com/ropensci/osmdata/issues/180

